---
title: "Stat331 Assignment 2"
author: "Milagro Chen (21066922)"
date: "`r Sys.Date()`"
output:
  pdf_document:
    keep_tex: yes
    latex_engine: xelatex
    number_sections: no
  html_document:
    toc: yes
  word_document: default
---

# Set up

Here are the libraries used:

```{r}
library(data.table)
library(tidyverse)
```

Here is also my A1 Question 2 code copy-pasted in, so that I can complete A2 Question 1 without hassle:
```{r}
#2.FROM A1 
truebeta <- c(-200000, 45, -300, 10000, 150000)
truesigma <- 12000
A1_variates <- read_table('~/Documents/GitHub/Big-Ahh-R-Repo/Stat331/Stat331 A1/data/raw/A1_variates (1).txt')
n <- nrow(A1_variates)
X <- as.matrix(cbind(rep(1,n),A1_variates))
set.seed(21066922)

#a)
simulatedresponse <- (as.vector(X %*% truebeta) + rnorm(n,0,truesigma))
#this is the simulated data with the random seed set to my student number
simulatedresponse

#b)
y <- simulatedresponse
size <- A1_variates$size
age <- A1_variates$age
employees <- A1_variates$employees
col <- A1_variates$col
mlr <- lm(y~size+age+employees+col)
summary(mlr)

#c)
betashat <- solve(t(X)%*%X)%*%t(X) %*% y
#d)
confint(mlr)
```

# FROM HERE ON IS THE START OF A2

# Question 1

Using the provided code, and my student number, I generated the betas:
```{r}
betahat_mat = array(0,c(5,1000)) #creates an array to store sets of estimates
set.seed(21066922)
for (i in 1:1000){
  y<-X%*%truebeta+rnorm(24,0,12000) #X and Beta vector as defined in A1.
  A1sim.lm<-lm(y~size+age+employees+col)
  betahat_mat[,i]<-coef(A1sim.lm) #yields estimates for each iteration
}
```
### a)
Then we will specifically focus on the beta_2:
```{r}
betahat_mat[2,]
```
and here is the histogram, where we can see it is roughly normal, as expected:
```{r}
hist(betahat_mat[2,])
```
We also expect that the mean be very close to 45, which is our true beta_2:
```{r}
mean(betahat_mat[2,])
```
We also also expect the standard deviation and the standard error to be very close:
```{r}
sqrt(var(betahat_mat[2,]))
```
Since we know that the beta estimate follows $N(truebeta, truesigma^2 (X^tX)^{-1})$,
we can get the standard error as such, confirming they are indeed very similar:
```{r}
sqrt((solve(t(X)%*%X) * truesigma**2)[2,2])
```
### b)
Focusing on the covariance now, we want to find the 3,4 entry of the covariance matrix:
```{r}
(solve(t(X)%*%X) * truesigma**2)[3,4]
```
Comparing it to our empirical(generated) result, we notice that it is double what is expected,
which just goes to show the randomness of the generated sample:
```{r}
cov(betahat_mat[3,],betahat_mat[4,])
```
# Question 2
So let's first load in the data, filter by positive revenue, and sample randomly with my student number as seed:
```{r}
ceocomp <- read.csv('~/Documents/GitHub/Big-Ahh-R-Repo/Stat331/Stat331 A2/data/raw/ceocomp.csv')
ceo_pos <- subset(ceocomp,ceocomp$PROF>0)
set.seed(21066922)
my.ceo_pos = ceo_pos[sample(nrow(ceo_pos),60),]   
my.ceo_pos 
```
### a)
Let's run the model:
```{r}
comp <- my.ceo_pos$COMP
x1 <- my.ceo_pos$AGE
x2 <- my.ceo_pos$EDUCATN
x3 <- as.factor(my.ceo_pos$BACKGRD)
x4 <- my.ceo_pos$TENURE
x5 <- my.ceo_pos$EXPER
x6 <- my.ceo_pos$SALES
x7 <- my.ceo_pos$VAL
x8 <- my.ceo_pos$PCNTOWN
x9 <- my.ceo_pos$PROF
fullmodel <- lm(comp ~ x1+x2+x3+x4+x5+x6+x7+x8+x9)
summary(fullmodel)
```
### b)
Let's run the hypothesis test for $\hat \beta = \vec 0$
```{r}
r2_full <- summary(fullmodel)$r.squared
msreg <- r2_full/12
msres <- (1-r2_full)/47
1 - pf(msreg/msres,12,47)
```
So with a significance level of 0.05, we can conclude with statistical significance that there is at least one explanatory variate that is non-zero, hence contributes to explaining the compensation.
### c)
```{r}
summary(fullmodel)
```
i) with beta hat x34 being -289, we say WITHOUT statistical significance(because the p value is 0.23>0.05) that lawyers tend to make $289 000 less than banking background, after factoring the rest of the explanatory variates.
ii) ceos with sales/marketing background also makes less, at -$56 000, than ones with banking background, accounting for all other explanatory factors. But we're even LESS certain of this estimate, with the p-value at 0.81.

Let's do an additional sum of squares test for the model without industry background as a factor, compared to the full model:
```{r}
redmodel <- lm(comp ~ x1+x2+x4+x5+x6+x7+x8+x9)
r2_red  <- summary(redmodel)$r.squared

f_num <- (r2_full - r2_red)/4
f_den <- (1 - r2_full)/47
f_stat <- f_num/f_den
1 - pf(f_stat,4,47)
```
Comparing it to the anova test, we find that the test comes back very similar:
```{r}
anova(redmodel,fullmodel)
```
Which means that after accounting for the other explantory variates, we can't conclude the CEO's background has much effect on the comp.
# d)
now let's focus on the reduced model:
```{r}
summary(redmodel)
```
Where we notice that the relation between age, tenure, and experience may be problematic because they are all correlated, as a long EXPER certainly implies longer TENURE and AGE.

Let's see the pairs chart:
```{r}
pairs(cbind(x1,x4,x5))
```
Where we see a slight linear relation between the 3 variables in question. So I'll say that multicollinearity may be an issue.
Let's test with the VIF for age:
```{r}
agemodel <- lm(x1~x2+x4+x5+x6+x7+x8+x9)
r2 <- summary(agemodel)$r.squared
VIF <- 1/(1-r2)
VIF
```
Which honestly is not that bad, so I retract my statement from before, as the VIF is fairly below 10. In case the assignment was looking for ONLY the 3 variables, here is that VIF:
```{r}
ageredmodel <- lm(x1~x4+x5)
r2 <- summary(ageredmodel)$r.squared
VIF <- 1/(1-r2)
VIF
```
Which it seems like it doesn't make much of a difference.
### e)
Let's predict a new CEO's comp:
```{r}
new_x <- data.frame(x1=48,x2=1,x4=14,x5=3,x6=3100,x7=2.1,x8=1.1,x9=172)
predict(redmodel,new_x,interval='prediction',level=.95)
```
We can say that the comp of 1.5 million is fairly reasonable, if not a little high. 1500 is still within the 95% prediction interval, so I'd say it is reasonable consistent with the other CEOs' comps.
