---
title: "Stat331 Assignment 3"
author: "Milagro Chen (21066922)"
date: "`r Sys.Date()`"
output:
  pdf_document:
    keep_tex: yes
    latex_engine: xelatex
    number_sections: no
  html_document:
    toc: yes
  word_document: default
---

# Set up

Here are the libraries used:

```{r}
library(data.table)
library(tidyverse)
library(leaps)
```

Here is also my A2 Question 2 code copy-pasted in, so that I can complete A3 without hassle:
```{r}
#2.FROM A2 
#a)
ceocomp <- read.csv('~/Documents/GitHub/Big-Ahh-R-Repo/Stat331/Stat331 A2/data/raw/ceocomp.csv')
ceo_pos <- subset(ceocomp,ceocomp$PROF>0)
set.seed(21066922)
my.ceo_pos = ceo_pos[sample(nrow(ceo_pos),60),]   
my.ceo_pos 
comp <- my.ceo_pos$COMP
x1 <- my.ceo_pos$AGE
x2 <- my.ceo_pos$EDUCATN
x3 <- as.factor(my.ceo_pos$BACKGRD)
x4 <- my.ceo_pos$TENURE
x5 <- my.ceo_pos$EXPER
x6 <- my.ceo_pos$SALES
x7 <- my.ceo_pos$VAL
x8 <- my.ceo_pos$PCNTOWN
x9 <- my.ceo_pos$PROF
fullmodel <- lm(comp ~ x1+x2+x3+x4+x5+x6+x7+x8+x9)
summary(fullmodel)

#b)
#keep in mind the hypothesis test is for betahat = 0 vector
r2_full <- summary(fullmodel)$r.squared
msreg <- r2_full/12
msres <- (1-r2_full)/47
1 - pf(msreg/msres,12,47)

#c)
#i) with beta hat x34 being -289, we say WITHOUT statistical significance that lawyers tend to make less than banking background
#ii) sales also makes less, but we're even LESS certain of this estimate.
#iii)
redmodel <- lm(comp ~ x1+x2+x4+x5+x6+x7+x8+x9)
r2_red  <- summary(redmodel)$r.squared

f_num <- (r2_full - r2_red)/4
f_den <- (1 - r2_full)/47
f_stat <- f_num/f_den
1 - pf(f_stat,4,47)
#iv)
anova(redmodel,fullmodel)
#d)
summary(redmodel)
#i) they are all correlated, as a long EXPER certainly implies longer TENURE and AGE
#ii) 
pairs(cbind(x1,x4,x5))
#iii)
agemodel <- lm(x1~x2+x4+x5+x6+x7+x8+x9)
summary(agemodel)
VIF <- 1/(1-0.2956)
#its ok
#e)
new_x <- data.frame(x1=48,x2=1,x4=14,x5=3,x6=3100,x7=2.1,x8=1.1,x9=172)
predict(redmodel,new_x,interval='prediction',level=.95)
```

# FROM HERE ON IS THE START OF A3

# Question 1

The residual vs fitted plot is fairly regular, nothing too out of the blue. Some outliers, but that's to be expected.

```{r}
plot(redmodel, which = 1)
```

The QQ is also rather normal, hinting at a slight right skew, but overall I think it's meets expectations of being roughly normally distributed.

```{r}
plot(redmodel, which = 2)
```

# Question 2

### a)

```{r}
hist(comp)
```

it has a heavy right tail, so using ln as a transform would send the outliers a lot closer to the rest of the data.

### b)

```{r}
lncomp <- log(comp)
lnredmodel <- lm(lncomp ~ x1+x2+x4+x5+x6+x7+x8+x9)
```

### c)

```{r}
summary(lnredmodel)
summary(redmodel)
```

we see a very slight improvement in the model in which we applied log to the compensation.

### d)

```{r}
plot(lnredmodel, which = 1)
```

I personally don't see a substantial difference in the res vs the fitted, but that's about normal for my human eyes.

```{r}
plot(lnredmodel, which = 2)
```

Here I notice that the skew is gone from before, but there is slightly lighter tails than expected of a normal distribution.

# Question 3

### a)

```{r}
plot(x6,lncomp)
```

Given the sales and log of compensation, it doesn't really seem like a linear model would be a good fit, the data doesn't seem to have a strong linear relation.

### b)

```{r}
plot(log(x6),lncomp)
```

We see a slightly better linear relation, but it's still not particularly clear, especially on the far right side.

### c)

```{r}
lnmodel <- lm(lncomp ~ log(x6)+log(x7)+log(x8)+log(x9))
summary(lnmodel)  
```

We can notice that the log of x7(Val), and log of x8(PCTOWN) have p values of significance, so we expect that these would explain more of the variation than the other 2 variates. Overall, it is a weaker model, with a smaller adj R^2 than the model previously, which only took the log of the response variate.

# Question 4

```{r}
plot(lnmodel, which = 1)
```

It is getting marginally better, with a flatter line, concentrating at 0.

```{r}
plot(lnmodel, which = 2)
```
The tails now are not as light as previous, so that's a bit more adequate.

# Question 5

```{r}
plot(fitted(lnmodel),rstudent(lnmodel))
abline(h=0)
```

We see that it is really similar to the previous plot.

```{r}
qqnorm(rstudent(lnmodel)); qqline(rstudent(lnmodel))
```
Same here, it is rather similar.

# Question 6

remember that p = 4, n = 60, so we consider 'high-leverage' to be:

```{r}
highlev = 2*(4+1)/60
highlev
```
```{r}
plot(hatvalues(lnmodel))
abline(h = highlev)
```
We see there is indeed a big outlier. Let's investigate.

# Question 7

Here is the highest leverage index:

```{r}
which.max(hatvalues(lnmodel))
```

Here are the plots of the explanatory variates with the 29th index highlighted:

```{r}
plot(log(x6))
points(29,log(x6)[29], col = "red", pch = 16, cex = 2.5)
```


```{r}
plot(log(x7))
points(29,log(x7)[29], col = "red", pch = 16, cex = 2.5)
```


```{r}

plot(log(x8))
points(29,log(x8)[29], col = "red", pch = 16, cex = 2.5)
```


```{r}
plot(log(x9))
points(29,log(x9)[29], col = "red", pch = 16, cex = 2.5)
```

I'd say that only in the log(x9), log of the profit, do we see index 29 being particularly outlier-ish.

# Question 8

```{r}
plot(lnmodel, which = 4)
```

We see that the 29th index is indeed the most influential, but according to our general rule in class, it isn't greater than 1, so we don't conclude it has 'high' influence, even though comparatively it has the highest influence.

# Question 9

### a)
let's start with the full model from question 3:

```{r}
summary(lnmodel)
```

Since log(x9) has the highest p value that is also above 0.15, we drop it:

```{r}
droppedx9 <- lm(lncomp ~ log(x6)+log(x7)+log(x8))
summary(droppedx9)
```

Since log(x6) has the highest p value above 0.15, we drop it:

```{r}
droppedx9x6 <- lm(lncomp ~ log(x7)+log(x8))
summary(droppedx9x6)
```

Since no other explanatory variate is above 0.15, we keep this as our model.

### b)

```{r}
lnX <- as.matrix(cbind(log(x6),log(x7),log(x8),log(x9)))
cp_out    <- leaps(x = lnX, y = lncomp, method = "Cp",    nbest = 1)
adjr2_out <- leaps(x = lnX, y = lncomp, method = "adjr2", nbest = 1)
cp_out$Cp    
cp_out$which  
adjr2_out$adjr2 
```

We pick log(x6), log(x7) and log(x8), bc its the highest adjR^2(barely though) that fits our mallows cp of being strictly less than k+1.

```{r}
pickedmodel <- lm(lncomp ~ log(x6)+log(x7)+log(x8))
```

### c)

```{r}
n <- length(lncomp)
mse_full <- summary(lnmodel)$sigma^2   
sse_reduced <- sum(resid(pickedmodel)^2)  
k <- 3+1                                

Cp_check <- sse_reduced/mse_full - (n - 2*k)
Cp_check

```

Which is the same mallows cp as expected.

### d) 

```{r}
summary(pickedmodel)
summary(droppedx9x6)
```

they are slightly different, but honestly its the exact same comparision made with the leaps function, with both models fitting our mallows cp criteria. Our picked model has 0.0001 more adj R^2 value, which is basically a negligible difference in my opinion.

### e)

```{r}
anova(pickedmodel,lnmodel)
```

From the test, we cant really say with confidence the dropped log(x9) variate, the log of the profit of the company, contributed significantly to the log of the CEO compensation.

### f)

```{r}
plot(fitted(pickedmodel), rstudent(pickedmodel))
abline(h = 0)
```

Once again, very similar res vs fitted values as previous.

```{r}
qqnorm(rstudent(pickedmodel)); qqline(rstudent(pickedmodel))
```

Very close to normal distribution, still with a little light of tails, though.

### g)

```{r}
new_lnx <- data.frame(x6=3100,x7=2.1,x8=1.1)
exp(predict(pickedmodel,new_lnx,interval='prediction',level=.95))
```

with a compensation of 1.5 million, we can now conclude that the ceo is being paid a little too much, given our model, and what we expected. As for the interval, it is tigher than the previous, as shown here:

```{r}
new_x <- data.frame(x1=48,x2=1,x4=14,x5=3,x6=3100,x7=2.1,x8=1.1,x9=172)
predict(redmodel,new_x,interval='prediction',level=.95)
```

So we went from a width of around 2000, to a width of around 1200, which is certainly a increase in certainty.










